---
title: "WUP_04"
author: "Anthony Falconio"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file = "GSScsv.csv")
```

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```
# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

I will be using chatGPT with CandP helper files to help create the code necessary to compare factor variables and numerical variables within the numerical and categorical data in the data set called GSS. We will then look at how WORDSUM results are effected by the level of education.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(data = mydata, aes(x= DEGREE, y = WORDSUM)) +
geom_boxplot ( notch=FALSE , outlier.color = "red" , outlier.size = 2 , fill ="darkgray") +
stat_boxplot( geom = 'errorbar' ) +
labs ( x = "DEGREE" , y = "WORDSUM SCORE" , title = "WORDSUM EXPLAINED BY DEGREE") +
theme ( plot.title = element_text (hjust = 0.5, size = 16))
```
The Boxplot indicates that the highest ranges, medians, and means for the WORDSUM scores are all among Graduate, Bachelor, and Junior College degrees. LT high school and high school degrees appear to have lower scores in regards to the WORDSUM test. There are some outliers with two graduate level degree scores being lower then its range, and one lt high school WORDSUM test result higher than the range of the lt high school degree in the data.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r include=TRUE,echo=FALSE}
favstats( WORDSUM ~ DEGREE , data=mydata)
```

The mean and median of the LT HIGH SCHOOL Degree are much lower than the median and mean of the other categories in regards to the results of the WORDSUM test results. The mean and median of the WORDSUM test results also increase in value in order from least to greatest from the lowest level of degree to the highest level degree within the dataset.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r include=TRUE, echo=FALSE}
model1 <- aov(WORDSUM ~ DEGREE, data = mydata)
summary(model1)
```
A P-value of less than 0.05 indicates that there is a significant difference in the data between people with varying levels of degrees with certain comparisons in regards to the WORDSUM test results.

```{r include=TRUE,echo=FALSE}
par(mar = c(3, 16, 2, 2))
plot(TukeyHSD(model1,conf.level = 0.99),las=2)
```

The TukeyHSD plot shows that there is a significant difference in the scores of the WORDSUM test between LT high school and a graduate level degree. There is also a significant difference between a school level degree and a graduate level degree. Both of these comparisons show a significant difference due to their intervals not containing the value of 0 as all the other group comparisons in the plot show. A 99% confidence level was also used to make sure that the results were stricter as a lower confidence level wouldn't have shown the lack of the value 0 in the intervals for lt high school and graduate as well as high school and graduate comparisons.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
The dataset includes the numerical variable WORDSUM and the character variable DEGREE which are compared to show a relationship between the two. The data shown from the boxplot shows that higher level degrees correlated with higher test scores for the WORDSUM test. This was also shown numerically with favstats as the mean and median of the degrees increased in value with each level of degree increasing. the one-way anova test showed that the p-value was <0.05 which showed a significant diferrence in the data between degrees so a TukeyHSD plot was used with a 99% confidence interval to see a significant difference between degree groups. There was a difference found in the WORDSUM scores between the less than a high school degree group and the graduate degree group. There was also a difference found in the WORDSUM scores between a high school degree and a graduate degree. There wasn't a significant difference among the other degree comparisons.
